---
title: 'P8105 Homework 5: Iteration'
author: "Annie Clark"
date: "November 9, 2018"
output: github_document
---

```{r}
library(tidyverse)
library(stringr)
library(purrr)
library(viridis)
```

# Problem 1
This zip file contains data from a longitudinal study that included a control 
arm and an experimental arm. Data for each participant is included in a 
separate file, and file names include the subject ID and arm.

## Part 1: Create a tidy dataframe

Create a tidy dataframe containing data from all participants, including the 
subject ID, arm, and observations over time:

* Start with a dataframe containing all file names; the list.files function will 
help
* Iterate over file names and read in data for each subject using purrr::map 
and saving the result as a new variable in the dataframe
* Tidy the result; manipulate file names to include control arm and subject ID, 
make sure weekly observations are “tidy”, and do any other tidying that’s 
necessary

```{r}
# List of file names
file_names = list.files("./data")

# Function to read cvs files and add file name
read_csv_function = function(file_name) {
  subject_data = 
    read_csv(file = str_c("./data/", file_name)) %>% 
    mutate(file = file_name)
}

## Load all files + tidy data
subject_data = 
  map_df(file_names, read_csv_function) %>% 
  gather(key = week, value = observation, week_1:week_8) %>% 
  separate(file, into = c("group", "subject"), sep = "_") %>% 
  mutate(week = str_replace(week, "week_", ""),
         subject = str_replace(subject,".csv",""),
         group = recode(group, "con" = "control", 
                        "exp" = "experimental"))

```


## Part 2: Make a plot

Make a spaghetti plot showing observations on each subject over time, and 
comment on differences between groups.


```{r}
subject_data %>% 
  mutate(group = recode(group, "control" = "Control", "experimental" = "Experimental")) %>% 
  ggplot(aes(x = week, y = observation, group = subject, color = subject)) + 
    geom_path() + 
    facet_grid(~group) + 
    labs(
      title = "Subject Observations over Time",
      x = "Week",
      y = "Observation"
    ) + 
    theme(legend.position = "bottom") +
    viridis::scale_color_viridis(
      name = "Subject ID", 
      discrete = TRUE
    )
```


# Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and 
made the data available through a GitHub repository here. You can read their 
accompanying article here.




```{r}
homicide_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

## Read in data from github
homicide_data = 
  read.csv(url(homicide_url)) %>% 
  janitor::clean_names() %>% 
  as_tibble()
```


## Part 1: Describe the data

### Prompt
Describe the raw data. 

### Response



## Part 2: Number of homocides and unsolved homocides per city

Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within 
cities to obtain the total number of homicides and the number of unsolved 
homicides (those for which the disposition is “Closed without arrest” or 
“Open/No arrest”).

```{r}
## Create city_state var and summarise by city_state
homicide_data =
  homicide_data %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  mutate(total_city_homicides = n()) %>% 
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) %>% 
  summarise(unsolved_homicides = n(),
            total_homicides = max(total_city_homicides))
```


## Part 3: Unsolved homocides in Baltimore

For the city of Baltimore, MD, use the prop.test function to estimate the 
proportion of homicides that are unsolved; save the output of prop.test as an 
R object, apply the broom::tidy to this object and pull the estimated proportion 
and confidence intervals from the resulting tidy dataframe.



```{r}
## Function to run prop_test + clean up results
homicide_prop = function(df) {
  homicide_prop_test = prop.test(df$unsolved_homicides, df$total_homicides, 
            alternative = "two.sided", conf.level = 0.95, correct = FALSE)
  
  broom::tidy(homicide_prop_test) %>% 
    select(estimate, conf.low, conf.high)
}

## Prop.test function applied to Baltimore, MD
homicide_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  homicide_prop() %>% 
  rename("Proportion of Unsolved Cases" = estimate, 
         "Lower bound (95% CI)" = conf.low, 
         "Upper bound (95% CI)" = conf.high) %>% 
  knitr::kable()
```



## Part 4: Unsolved homocides per city

Now run prop.test for each of the cities in your dataset, and extract both the 
proportion of unsolved homicides and the confidence interval for each. Do this 
within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns 
and  unnest as necessary to create a tidy dataframe with estimated proportions 
and CIs for each city.


```{r}
## Prop test function applied to all cities
homicide_data = 
  homicide_data %>% 
  nest(unsolved_homicides:total_homicides) %>% 
  mutate(prop_tests = map(.x = .$data, ~homicide_prop(.x))) %>% 
  unnest() %>% 
  rename(unsolved_estiamte = estimate, lower_bound = conf.low,
         upper_bound = conf.high)
```

## Part 5: Create a plot

Create a plot that shows the estimates and CIs for each city – check out 
geom_errorbar for a way to add error bars based on the upper and lower limits. 
Organize cities according to the proportion of unsolved homicides.


```{r}
homicide_data %>% 
  mutate(city_state = forcats::fct_reorder(city_state, desc(unsolved_estiamte))) %>%
  ggplot(aes(x = city_state, y = unsolved_estiamte, color = city_state)) + 
    geom_point() +
    geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound, width = 0.2)) + 
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    labs(
        title = "Proportion of Unsolved Homocides in 50 Large Cities",
        x = "City, State",
        y = "Proportion of Unsolved Cases", 
        caption = "Data from the Washington Post"
      ) +
    theme(legend.position = "none") +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

