---
title: 'P8105 Homework 5: Iteration'
author: "Annie Clark"
date: "November 9, 2018"
output: github_document
---

```{r}
library(tidyverse)
library(stringr)
library(purrr)
library(viridis)
```

# Problem 1
This zip file contains data from a longitudinal study that included a control 
arm and an experimental arm. Data for each participant is included in a 
separate file, and file names include the subject ID and arm.

## Part 1: Create a tidy dataframe

Create a tidy dataframe containing data from all participants, including the 
subject ID, arm, and observations over time:

* Start with a dataframe containing all file names; the list.files function will 
help
* Iterate over file names and read in data for each subject using purrr::map 
and saving the result as a new variable in the dataframe
* Tidy the result; manipulate file names to include control arm and subject ID, 
make sure weekly observations are “tidy”, and do any other tidying that’s 
necessary

```{r}
## List of file names
file_names = list.files("./data")

## Function to read cvs files and add file name
read_csv_function = function(file_name) {
  subject_data = 
    read_csv(file = str_c("./data/", file_name)) %>% 
    mutate(file = file_name)
}

## Load all files + tidy data
subject_data = 
  map_df(file_names, read_csv_function) %>% 
  gather(key = week, value = observation, week_1:week_8) %>% 
  separate(file, into = c("group", "subject"), sep = "_") %>% 
  mutate(week = str_replace(week, "week_", ""),
         subject = str_replace(subject,".csv",""),
         group = recode(group, "con" = "control", 
                        "exp" = "experimental"))

?fct_recode
```


## Part 2: Make a plot

Make a spaghetti plot showing observations on each subject over time, and 
comment on differences between groups.

```{r}
subject_data %>% 
  mutate(group = recode(group, "control" = "Control", "experimental" = "Experimental")) %>% 
  ggplot(aes(x = week, y = observation, group = subject, color = subject)) + 
    geom_path() + 
    facet_grid(~group) + 
    labs(
      title = "Subject Observations over Time",
      x = "Week",
      y = "Observation"
    ) + 
    theme(legend.position = "bottom") +
    viridis::scale_color_viridis(
      name = "Subject ID", 
      discrete = TRUE
    )
```

The observations for the subjects in the control arm over the 8 weeks stay relatively constant over (and potentially decreasing), while the observations for the subjects in the experimental arm have an increasing trend. I would be interested in doing additional statistical analysis to confirm that the treatment had a significant effect on the outcome. 

Please note: it is unclear if the subject IDs are unique within the control/experimental arms or are global subject IDs (suggesting that each of the IDs went through both the control and experimental arms). If the subject IDs are only unique within the arms, I would change the above graphs to make it clearer that the subjects with the same subject IDs from both arms are different.  

# Problem 2
The Washington Post has gathered data on homicides in 50 large U.S. cities and 
made the data available through a GitHub repository here. You can read their 
accompanying article here.


```{r}
homicide_url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"

## Read in data from github
homicide_data = 
  read.csv(url(homicide_url)) %>% 
  janitor::clean_names() %>% 
  as_tibble()
```


## Part 1: Describe the data

### Prompt
Describe the raw data. 

### Response
The homicide dataset has `r nrow(homicide_data)` rows and `r ncol(homicide_data)` columns. Each row contains information about a single homicide that has occurred, such as location (city, state, latitude/longitude), victim information (first/last name, age, sex, disposition) and information about the homicide (disposition, reported date, id). The date range of reports spans from `r min(homicide_data$reported_date)` to 20171231. There are two invalid reported date values that are most likely typos ("201511105" and "201511018"). The data includes `r count(homicide_data, city) %>% nrow()` cities in `r count(homicide_data, state) %>% nrow()` states. There are a total of 51 distinct city/state combinations, including Tulsa, AL. This is most likely a typo for Tulsa OK (confirmed with the longitude/latitude coordinates). Because this particular homicide was closed with an arrest and the primary purpose of the analysis is to look at unsolved homicides, this case was not included in the analysis. 

`r round(homicide_data %>% filter(disposition == "Closed by arrest") %>% nrow() / nrow(homicide_data) * 100)`% of the homicides were solved by arrest and `r round(homicide_data %>% filter(disposition != "Closed by arrest") %>% nrow() / nrow(homicide_data) * 100)`% of the cases are considered unsolved (solved with no arrest or still open/no arrest). 

`r round(homicide_data %>% filter(victim_sex == "Male") %>% nrow() / nrow(homicide_data) * 100)`% of the homicides had male victims, `r round(homicide_data %>% filter(victim_sex == "Female") %>% nrow() / nrow(homicide_data) * 100)`% of homicides had female victims and the remaining `r round(homicide_data %>% filter(victim_sex == "Male") %>% nrow() / nrow(homicide_data) * 100)`% of cases had unknown genders of the victims. 


## Part 2: Number of homicides and unsolved homicides per city

Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within 
cities to obtain the total number of homicides and the number of unsolved 
homicides (those for which the disposition is “Closed without arrest” or 
“Open/No arrest”).

```{r}
## Create city_state var and summarise by city_state
homicide_data =
  homicide_data %>% 
  mutate(city_state = str_c(city, state, sep = ", ")) %>% 
  group_by(city_state) %>% 
  mutate(total_city_homicides = n()) %>% 
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) %>% 
  summarise(unsolved_homicides = n(),
            total_homicides = max(total_city_homicides))
```


## Part 3: Unsolved homicides in Baltimore

For the city of Baltimore, MD, use the prop.test function to estimate the 
proportion of homicides that are unsolved; save the output of prop.test as an 
R object, apply the broom::tidy to this object and pull the estimated proportion 
and confidence intervals from the resulting tidy dataframe.



```{r}
## Function to run prop_test + clean up results
homicide_prop = function(df) {
  homicide_prop_test = prop.test(df$unsolved_homicides, df$total_homicides, 
            alternative = "two.sided", conf.level = 0.95, correct = FALSE)
  
  broom::tidy(homicide_prop_test) %>% 
    select(estimate, conf.low, conf.high)
}

## Prop.test function applied to Baltimore, MD
homicide_data %>% 
  filter(city_state == "Baltimore, MD") %>% 
  homicide_prop() %>% 
  rename("Proportion of Unsolved Cases" = estimate, 
         "Lower bound (95% CI)" = conf.low, 
         "Upper bound (95% CI)" = conf.high) %>% 
  knitr::kable()
```



## Part 4: Unsolved homicides per city

Now run prop.test for each of the cities in your dataset, and extract both the 
proportion of unsolved homicides and the confidence interval for each. Do this 
within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns 
and  unnest as necessary to create a tidy dataframe with estimated proportions 
and CIs for each city.


```{r}
## Prop test function applied to all cities
homicide_data = 
  homicide_data %>% 
  nest(unsolved_homicides:total_homicides) %>% 
  mutate(prop_tests = map(.x = .$data, ~homicide_prop(.x))) %>% 
  unnest() %>% 
  rename(unsolved_estiamte = estimate, lower_bound = conf.low,
         upper_bound = conf.high)
```

## Part 5: Create a plot

Create a plot that shows the estimates and CIs for each city – check out 
geom_errorbar for a way to add error bars based on the upper and lower limits. 
Organize cities according to the proportion of unsolved homicides.


```{r}
homicide_data %>% 
  mutate(city_state = forcats::fct_reorder(city_state, desc(unsolved_estiamte))) %>%
  ggplot(aes(x = city_state, y = unsolved_estiamte, color = city_state)) + 
    geom_point() +
    geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound, width = 0.2)) + 
    theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
    labs(
        title = "Proportion of Unsolved Homicides in 50 Large Cities",
        x = "City, State",
        y = "Proportion of Unsolved Cases", 
        caption = "Data from the Washington Post"
      ) +
    theme(legend.position = "none") +
    viridis::scale_color_viridis(
      discrete = TRUE
    )
```

The city with the largest proportion of unsolved homicides is Chicago (0.74) and it also has very narrow bounds on the 95% confidence interval for the point estimate. This means that there is more certainty around the number of unsolved homicides in Chicago being over 0.7. The lowest proportion of unsolved homicides is Richmond Virginia. The cities with the highest variability in unsolved homocide proportions are Savannah, GA and Tampa, FL. 